Parsing Postfix log files

### 2008-04-01 doc/logparser.tex
Corrections from Carl for sections 3 and 4.

            \item [POSTFIX\_RELOAD] The administrator has caused Postfix to
                start or stop, and all existing \daemon{smtpd} processes
                will be terminated.

@@

    \item [result\_data, connection\_data] Sometimes rules need to supply a
        piece of data which isn't present in the log line: e.g.\ setting
        \texttt{smtp\_code} when mail is accepted.  The format and allowed
        variables are the same as for \texttt{result\_cols} and
        \texttt{connection\_cols}, except that arbitrary
        data\footnote{Commas and semi-colons cannot be escaped and thus
        cannot be used.  This is intended for use with small amounts of
        data rather than large amounts, so dealing with escape sequences
        seemed unnecessary.} is permitted on the right hand side of the
        assignment.

    \item [action] The action that will be invoked when this rule matches a
        log line; a full list can be found in
        \sectionref{actions-in-detail}.

    \item [queueid] Specifies the matching capture from the \regex{} which
        gives the queueid, or zero if the log line doesn't contain a
        queueid.

@@

    \item [start] The timestamp of the first log line, in seconds since the
        epoch.

@@

    \item [sender] The sender's email address.  This can change during one
        single connection, when the connection is reused to send multiple
        mails.

@@

    \item There are no sub-rules, so rules have linear computational
        complexity.

@@

This rule would match the following log line:

@@

An important property of a parser is how execution time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
Graph~\refwithpage{execution time vs file size vs number of lines graph}
shows the execution time in seconds, file size in MB and tens of thousands
of log lines per log file.  All three lines run roughly in parallel, giving
a visual impression that the algorithm scales linearly with input size.
This impression is borne out by graph~\refwithpage{execution time vs file
size vs number lines factor}, which plots the ratio of file size vs
execution time and ratio of number of log lines vs execution time (higher
is better).\footnote{Table~\refwithpage{execution time vs file size vs
number lines factor table} shows the ratios for different breakdowns of the
log files.}  As the reader can see the ratios are quite tightly banded,
showing that the algorithm scales linearly: the much larger log files
between points 60 and 70 on the X axis in graph~\refwithpage{execution time
vs file size vs number of lines graph} actually cause the ratio to
increase, rather than decrease.  The strange behaviour where larger log
files are parsed more efficiently is explained fully in \sectionref{Why are
there dips in the graphs?}.

>>>>

            \item [POSTFIX\_RELOAD] The administrator has instructed
                Postfix to start or stop, and all existing \daemon{smtpd}
                processes will be terminated.  This does not negatively
                impact on the logs or mail queued by Postfix for delivery.

@@

    \item [result\_data, connection\_data] Sometimes rules need to supply a
        piece of data which isn't present in the log line: e.g.\ setting
        \texttt{smtp\_code} when mail is accepted.  The format and allowed
        variables are the same as for \texttt{result\_cols} and
        \texttt{connection\_cols}, except that arbitrary
        data\footnote{Commas and semi-colons cannot be escaped and thus
        cannot be used.  This is intended for use with small amounts of
        data rather than large amounts in any one rule, so dealing with
        escape sequences was deemed unnecessary.} is permitted on the right
        hand side of the assignment.

    \item [action] The action that will be invoked when this rule matches a
        log line; a full list of actions and the parameters they are
        invoked with can be found in \sectionref{actions-in-detail}.

    \item [queueid] Specifies the matching capture from the \regex{} which
        gives the queueid, or zero if the log line doesn't contain a
        queueid.  Many log lines won't contain a queueid, e.g.\ rejections
        logged before a mail has been accepted (a queueid won't have been
        allocated), or log lines which aren't tied to one particular mail.

@@

    \item [start] The timestamp of the first log line, in seconds since the
        epoch (explained in the glossary, appendix~\refwithpage{Glossary}).

@@

    \item [sender] The sender's email address.  This can change during one
        single connection, when the connection is reused to send multiple
        mails.  XXX BETTER EXPLANATION\@.

@@

    \item Rule processing time is a linear function of the number of rules.

@@

The example rule below would match the following log line:

@@

An important property of a parser is how execution time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
Graph~\refwithpage{execution time vs file size vs number of lines graph}
shows the execution time in seconds, file size in MB and tens of thousands
of log lines per log file.  All three lines run roughly in parallel, giving
a visual impression that the algorithm scales linearly with input size.
This impression is borne out by graph~\refwithpage{execution time vs file
size vs number lines factor} which plots the ratio of file size vs
execution time and ratio of number of log lines vs execution time (higher
is better).\footnote{Table~\refwithpage{execution time vs file size vs
number lines factor table} shows the ratios for different breakdowns of the
log files.}  As the reader can see the ratios are quite tightly banded,
showing that the algorithm scales linearly: the much larger log files
between points 60 and 70 on the X axis in graph~\refwithpage{execution time
vs file size vs number of lines graph} actually cause the ratio to
increase, rather than decrease.  The strange behaviour where larger log
files are parsed more efficiently is explained fully in \sectionref{Why are
there dips in the graphs?}.
